The current prototype is strongest as a research framework: it exposes the full chain from OSM public GPS signals to route reconstruction, feature extraction, profile construction, candidate generation, ranking, and evaluation. The GUI/demo angle remains valuable because it shows how a user would interact with ranked routes and profile explanations.

The main limitation is data validity. OSM public GPS trackpoints cannot be assumed to represent clean, persistent user histories. The current pseudo-history profiles are useful for exploring whether trace-derived movement signals can inform route ranking, but they are not equivalent to logged user route choices. This affects both profile construction and evaluation.

The second limitation is reconstruction quality. GPS-to-route distance is low, but route-distance ratio is high for several segments. The threshold-sensitivity study shows this limitation persists across reasonable segmentation and filtering choices. The Porto benchmark now uses simplified GPS anchors before shortest-path stitching, which modestly reduces route-distance-ratio inflation, but it does not solve the problem. This suggests the approximate nearest-node plus shortest-path stitching method can still overbuild routes. Before submission, the pipeline should improve map matching, segment splitting, or route simplification so strict usability passes without relaxing the route-distance-ratio threshold.

The third limitation is evaluation size and external comparability. The current route-candidate evaluation uses eight held-out queries from one OSM public trace area and 50 sampled Porto queries. This is enough for a transparent prototype and benchmark framework, but not enough for a claim that the method beats prior personalized route-recommendation systems.

The main path to a strong paper is a same-data public benchmark. NASR-style evaluation on a public dataset such as Porto taxi trajectories would let us compare route-recovery or route-ranking methods with recognized metrics such as precision, recall, F1, and edit distance. A direct claim of superiority should only be made after matching the dataset, task definition, train/test split, and metrics of an external baseline or a faithful reproduction.

The Porto prototype is now useful mainly as a diagnostic. It shows that the vehicle-aware profile scorer can be evaluated on public trajectory data and improves over shortest-distance and the generic profile on several ranking metrics. The diverse candidate generator also improves the path-oracle ceiling relative to the earlier five-candidate setup. However, random ranking remains close on some top-$k$ and path metrics, and the oracle rows still show that path-level quality is constrained by the generated candidate pool and reconstructed route references. This points to the next research problem: the candidate-generation/oracle setup and map matching must be strengthened, and the benchmark must be expanded before it can support a strong publication claim.

Prompt/SBERT and hybrid results currently use synthetic preference text generated from reconstructed route features. These are useful as ablations because they test the prompt-ranking machinery, but they do not prove performance with real user-stated preferences. If prompt or hybrid personalization becomes a central claim, the system needs collected preference text or a controlled preference study.

\subsection{Threats to Validity}
The most important internal-validity threat is oracle definition. Feature-oracle ranking rewards closeness to reconstructed route features, which may favor methods that match aggregate properties rather than exact geometry. Path-oracle metrics partly address this, but they remain bounded by reconstruction quality. The most important external-validity threat is that walking-style OSM pseudo-histories and Porto taxi trajectories represent different travel modes. We therefore do not combine them into one headline result. The most important construct-validity threat is interpreting pseudo-history profiles as user preferences. The paper avoids that interpretation unless clean user identity or stated preferences are available.

\subsection{Path to a Stronger Claim}
The next publishable claim should be tested on a larger same-data benchmark. A defensible target is: given public Porto trajectories and OSM road networks, compare profile-aware candidate reranking against shortest-path, random, learned feature ranking, and a reproduced NASR-style route-recovery baseline using the same path-overlap metrics. If the profile-aware method improves ranking metrics while maintaining competitive path F1/NDTW and the GUI/API remains usable, the project can claim an interpretable, open-data route-ranking alternative. If it does not, the benchmark will still identify whether the limiting factor is profile scoring, candidate generation, or map matching.
