\subsection{Personalized Route Recommendation}
Personalized route recommendation has been studied with stronger data assumptions than this work can currently make. Dai et al. use large-scale trajectory data to model personalized route recommendation from historical movement behavior~\cite{dai2015bigtrajectory}. More recent user-habit work models route sorting from historical navigation behavior and compares against minimum-ETA, LightGBM, and DCN-v2 baselines with inconsistency-oriented metrics~\cite{userhabits2024}. These papers motivate our use of route features and historical profiles, but they assume cleaner user or vehicle histories than OSM public trace signals provide.

Neural search methods frame route recommendation as an origin-destination search problem. NASR integrates neural models with A* search for personalized route recommendation and evaluates route recovery with path-level metrics such as precision, recall, F1, and edit distance~\cite{nasr2019}. This is the strongest top-tier methodological reference for our next benchmark target because it affects route generation itself rather than only post-hoc reranking. Our current system is simpler: it generates a small set of OSM candidates and reranks them with interpretable profile or prompt signals. The Porto benchmark is designed to move the evaluation closer to NASR-style route recovery without pretending that a full NASR reproduction has already been completed.

\subsection{Trajectory Reconstruction and Historical Route Planning}
Trajectory reconstruction and popular-route mining are directly relevant because our pseudo-history profiles depend on reconstructed traces. Chen et al. study popular-route construction from uncertain trajectories and use trajectory similarity metrics such as normalized dynamic time warping and maximum distance~\cite{chen2012popular}. Tian et al. propose DRPK for effective route planning from historical trajectories on road networks~\cite{tian2023drpk}. These works motivate a strict separation between reconstruction quality and ranking quality. A route ranker can appear to perform well if the reconstructed reference path is poor or if the candidate pool is artificially easy; therefore, we report route-distance ratio, GPS-to-route distance, path F1, NDTW, edit distance, and oracle upper bounds.

\subsection{Contextual and Preference-Aware Recommendation}
Contextual route recommendation datasets and route-choice studies show that origin-destination routing is affected by context, not only graph distance~\cite{contexttrails2025}. Preference-driven travel-planning systems such as Personal Travel Solver use LLM and solver components to infer user preferences and produce personalized travel plans~\cite{wang2025pts}. Our work is narrower. We do not claim full itinerary planning or language-grounded route generation. Instead, natural-language preferences are optional inputs used to rerank already-generated OSM route candidates. When preference text is synthetic, we label it as synthetic and treat prompt/SBERT results as ablations rather than evidence of real stated preferences.

\subsection{Positioning}
The closest direct future comparison is not the OSM public-trace experiment alone. The defensible comparison path is a same-data public benchmark, most likely Porto, where our candidate reranker and external route-recovery baselines are evaluated with the same path-overlap metrics. Until that reproduction is complete, this paper positions the proposed method as an interpretable, open-data route-ranking framework with transparent diagnostics, rather than as a system that already surpasses learned personalized search methods.

\begin{table*}[t]
  \centering
  \caption{Baseline alignment. The table separates conceptual relevance from direct comparability so that the evaluation does not force an invalid comparison.}
  \label{tab:baseline_alignment}
  \begin{tabular}{p{0.22\textwidth}p{0.19\textwidth}p{0.17\textwidth}p{0.17\textwidth}p{0.17\textwidth}}
    \toprule
    Work & Main task & Data signal & Comparable metrics & Role in this paper \\
    \midrule
    User-habit route recommendation~\cite{userhabits2024} & Personalized vehicle route sorting & Historical navigation logs & Inconsistency-oriented ranking metrics & Closest conceptual profile-ranking reference, but not directly comparable without clean user logs. \\
    NASR~\cite{nasr2019} & Personalized route search/recovery & Trajectory data and road graphs & Precision, recall, F1, edit distance & Main future technical benchmark target on public trajectory data. \\
    Big trajectory PRR~\cite{dai2015bigtrajectory} & Driver-specific route recommendation & Large GPS trajectory data & Route-cost and preference-oriented evaluation & Older but important driver-preference baseline family. \\
    RICK~\cite{chen2012popular} & Popular route inference from uncertain trajectories & Sparse/noisy trajectories & NDTW, maximum distance & Reconstruction-quality reference, not a personalized ranking baseline. \\
    DRPK~\cite{tian2023drpk} & Realistic route planning from historical trajectories & Large historical trajectories & Path prediction and efficiency metrics & Non-personalized historical-trajectory planning reference. \\
    Context Trails / PTS~\cite{contexttrails2025,wang2025pts} & Contextual route/travel recommendation & Contextual POI trails or reviews & Recommendation and preference-satisfaction metrics & Motivates context and preference modeling; task differs from OSM route reranking. \\
    \bottomrule
  \end{tabular}
\end{table*}
